---
title: "RWorksheet_Olivo#4c"
author: "Mel Adry Olivo"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
Download and open the mpg file. Upload it to your OWN environment

1a. Show your solutions on how to import a csv file into the environment.

```{r}
library(readr)

mpg_file <- read.csv("mpg.csv")
mpg_file

```

1b.Which variables from mpg dataset are categorical?

```{r}

str(mpg_file)

# manufacturer, model, trans, drv, fl, class variables are categorical

```

1c. Which are continuous variables?

```{r}
str(mpg_file)

# X, displ, year, cyl, cty, hwy are continuous variables
```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}

manufacturer_asTable <- table(mpg_file$manufacturer)
names(manufacturer_asTable)[which.max(manufacturer_asTable)]

# dodge manufacturer has the most models

model_asTable <- table(mpg_file$model)
model_asTable
names(model_asTable)[which.max(model_asTable)]

# caravan 2wd has the most variations

```


2a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}

manufacturers_models <- data.frame(Manufacturer = mpg_file$manufacturer, Model = mpg_file$model)

unique_mods <- unique(manufacturers_models)

unique_mods
unique_mods$mpg_file.model

```

2b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}


```