---
title: "RWorksheet_Olivo#4b"
author: "Mel Adry Olivo"
date: "2023-11-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. 

It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r name, echo = TRUE}
matZero <- matrix(c(0,0,0,0,0), 5, 5)
vectorA <- c(1,2,3,4,5)

for (i in 1:length(vectorA)) {
  matVec0[i, ] <- abs(vectorA - vectorA[i])
}
print(matVec0)

```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure 2

```{r name, echo = TRUE}
for(i in 1:5) {
  stars <- rep("*", i)
  print(stars)
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r name, echo = TRUE}

userInput <- as.integer(readline("Enter starting number for Fibonacci sequence: "))

x <- userInput
y <- 0


cat("Fibonacci sequence starting from", userInput, ":\n")

repeat {
  
  next_num <- x + y
  
  if (next_num > 500){
    break
  }
  cat(next_num, " ")
  x <- y
  y <- next_num
  
}

```

4. Import the dataset as shown in Figure 1 you have created previously.

4a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? 
Show your codes and its result.

```{r name, echo = TRUE}
prevData <- read.csv("householdData.csv")
head(prevData)
```

4b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? 
Write the R scripts and its output.

```{r name, echo = TRUE}
males <- prevData[prevData$Gender == "M",]
males

females <- prevData[prevData$Gender == "F",]
females

numofMale <- nrow(males)
numofMale

numofFem <- nrow(females)
numofFem
```


4c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. 
Write the R scripts and its result

```{r name, echo = TRUE}

totalMaleFemale <- table(prevData$Gender)
barplot(totalMaleFemale,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("red", "blue"),
        legend.text = rownames(totalMaleFemale),
        beside = TRUE)
```

5. The monthly income of Dela Cruz family was spent on the following:

Food Electricity Savings Miscellaneous
60      10         5         25

5a. a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.

```{r name, echo = TRUE}

spending_data <- data.frame(
  Category = c("Food", "Electricity", "Savings", "Miscellaneous"),
  Value = c(60, 10, 5, 25)
)

spending_data$Percentage <- spending_data$Value / sum(spending_data$Value) * 100

colors <- c("red", "blue", "green", "purple")

pie(spending_data$Value,
    labels = paste(spending_data$Category, "(",spending_data$Percentage,"%)"),
    col = colors,
    main = "Monthly Income Spending of Dela Cruz Family")

legend("topright", spending_data$Category, fill = colors)


```
